---
title: "8: STAT 213 Review"
format:
  html:
    embed-resources: true
    warning: false
    message: false
documentclass: article
author: "Bryana A. S. Thieret"
mainfont: "Cambria"
monofont: "Cambria"
---

## 8.1: STAT 213 with `broom`  

* In the examples below, we will consider the multiple regression model  

$$Y = \beta_0 + \beta_1species + \beta_2aroma + \beta_3 flavor + \beta_4sweetness + \beta_5moisture + \epsilon$$  

* where $Y$ is the rating of the coffee (total_cup_points), species in an indicator variable equal to 1 if the species is Robusta and a 0 if the species is Arabica.  

### 8.1.1: `tidy()`  

* `tidy()` is analogous to `summary()` for a linear model.  

```{r}
library(tidyverse)
library(broom)
library(here)
theme_set(theme_classic())

coffee_df <- read_csv(here("Section Notes/data/coffee_ratings.csv"))


coffee_mod <- lm(total_cup_points ~ species + aroma + flavor +
                   sweetness + moisture,
   data = coffee_df)
```

```{r}
summary(coffee_mod)

tidy(coffee_mod)
```

* The advantage of this format of output is that we can now use other `tidyverse` functions on the output.  

```{r}
# to pull p-values 
tidy(coffee_mod) |> select(p.value)

# grab the output for a particular variable of interest
tidy(coffee_mod) |> filter(term == "aroma")
```


### 8.1.2: `glance()`  

* `glance()` puts some model summary statistics into a tidy tibble. 

```{r}
glance(coffee_mod)
```


### 8.1.3: `augment()`  

* `augment()` returns a tibble that contains all of the variables used to fit the model appended with commonly used diagnostic statistics like the fitted values, cooks distance, values for leverage, and residuals.  

```{r}
augment(coffee_mod)

# filter() the data set to examine values with high cook's distance that might be influential

augment_df = augment(coffee_mod)
augment_df |> filter(.cooksd > 1)

augment_df |> filter(.fitted > 25) |> 
  ggplot(aes(x = .fitted, y = .resid)) + 
  geom_point()
```

### 8.1.4: Exercises  

**Exercise 1:** Examine the penguins data set in the palmerpenguins package. Fit a linear regression model with body_mass_g as the response variable and species and bill_length_mm as the predictors. Note that penguins with missing values for any of these three variables will be dropped from the analysis.  

```{r}
library(palmerpenguins)
penguins

pen_mod1 = lm(body_mass_g ~ bill_length_mm, data = penguins)
```



**Exercise 2:** Create a table of summary output, including coefficient estimates, standard errors, test statistics, and p-values, using one of the broom functions.

```{r}
tidy(pen_mod1)
```


**Exercise 3:** Use glance() to glance at some of the relevant model statistics.  

```{r}
glance(pen_mod1)
```


**Exercise 4:** Using augment(), create a plot of the residuals vs. the fitted values and evaluate the constant variance assumption.  

```{r}
augment(pen_mod1) |> ggplot(aes(x = .fitted, y = .resid)) + 
  geom_point() +
  geom_hline(yintercept = 0.5, color = "red")
```


**Exercise 5:** Using augment(), check to see if there are any penguins that are influential. Use 0.75 as your cut-off value for Cook’s distance.


```{r}
augment(pen_mod1) |> filter(.cooksd > .75)
```
* There are no penguins that are particularly influential in this data set.  

_-----------------------------------------------------------------------------------------_  

## 8.2: Model Coefficient Interpretation  

```{r}
library(here)
library(tidyverse)
library(broom)

statsurvey_df <- read_csv(here("Section Notes/data/stat113_survey.csv"))

statsurvey_nomiss <- statsurvey_df |> filter(!is.na(GPA))
statsurvey_nomiss <- statsurvey_nomiss |>
  mutate(time_year = as.numeric(time_year)) |>
  mutate(a_year = if_else(time_semester == "S",
                          true = time_year - 1,
                          false = time_year)) |>
  filter(GPA <= 4.0) |>
  mutate(a_year_fact = as.factor(a_year),
         a_year = a_year + 2000)
```


```{r}
year_linear = lm(GPA ~ a_year, data = statsurvey_nomiss)
tidy(year_linear)
```

```{r}
yearlm = lm(GPA ~ a_year_fact, data = statsurvey_nomiss)
tidy(yearlm)
```

```{r}
glance(year_linear)
glance(yearlm)
anova(year_linear, yearlm)
```


```{r}
# creates "grid"
grid = tibble::tibble(a_year = 2005:2021)
auglinear = augment(year_linear, newdata = grid, interval = "confidence")

auglinear |> ggplot(aes(y = .fitted, x = a_year)) + 
  geom_jitter(data = statsurvey_nomiss, aes(x = a_year, y = GPA), alpha = 0.2, shape = 21) +
  geom_point(color = "purple") +
  geom_errorbar(aes(ymin = .lower, ymax = .upper), width = 0.25, color = "purple") +
  labs(caption = "Error bars are 95% CI for the mean response of the fitted value.") +
  theme_classic() 
```

Check Assumptions

* qqnorm 
* residual vs. fitted 
* 
```{r}
yearlm
year_aug <- yearlm |> augment()

# linearity 
## scatter plot of residual vs fitted 
year_aug |> ggplot(aes(y = .resid, x = .fitted)) + geom_point()

# normality 
## histogram of residuals
year_aug |> ggplot(aes(x = .resid)) + geom_histogram(fill = "slateblue", color = "blue")


# equal variance 
year_aug |> ggplot(aes(x = .fitted, y = .resid)) + 
  geom_jitter(width = 2.5, alpha = 0.6)



```
















